Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-06 ms.tex
abstract work

\begin{abstract}
We calculate steady-state, one-dimensional hydrodynamic profiles of hot gas in the nuclei of nominally quiescent galaxies for a range of central black hole mass $M_{\bullet}$, parameterized gas heating rate, and observationally-motivated stellar density profiles.  Mass is supplied to the circumnuclear medium by stellar winds, while gas is heated primarily by stellar winds, supernovae, and black hole feedback.  Analytic estimates are provided for the stagnation radius (where the radial velocity of the gas passes through zero) and the black hole mass accretion rate, $\dot{M}$, as a function of black hole mass and the gas heating efficiency, the latter of which is related to the star-formation history.  We assess the conditions under which radiative instabilities develop, separately in the case of a single burst of star formation and in the case of the average star formation history of galaxies hosting a given black hole mass. The nuclear X-ray luminosities of a sample of quiescent black holes are combined with our results for $\dot{M}$ and an estimate of the radiative efficiency of the accretion flow to conclude that quiescent black holes in the local universe are consistent with accreting in a locally thermally-stable state.  The maximum accretion rate of a thermally-stable flow is, however, too low to explain the accretion rates and short growth time of low mass black holes in the local universe.  The latter must instead result from gas being fed in from large radii, either due to galaxy mergers associated with hierarchical growth or thermal instabilities on larger (galactic) scales.  We briefly discuss implications of our results for the use of an assumed $\dot{M}-M_{\bullet}$ relationship to constrain the occupation fraction of SMBHs in low mass galaxies and the diversity of circumnuclear environments encountered by the jetted outflows from tidal disruption events.
\end{abstract}

>>>>

\begin{abstract}
We calculate steady-state, one-dimensional hydrodynamic profiles of hot gas in the nuclei of nominally quiescent galaxies for a range of central black hole mass $M_{\bullet}$, parameterized gas heating rate, and observationally-motivated stellar density profiles.  Mass is supplied to the circumnuclear medium by stellar winds, while gas is heated primarily by stellar winds, supernovae, and black hole feedback.  Analytic estimates are provided for the stagnation radius (where the radial velocity of the gas passes through zero) and the black hole mass accretion rate, $\dot{M}$, as a function of black hole mass and the gas heating efficiency, the latter of which is related to the star-formation history.  We assess the conditions under which radiative instabilities develop, separately in the case of a single burst of star formation and in the case of the average star formation history of galaxies hosting a given black hole mass. The nuclear X-ray luminosities, $L_{X}$, of a sample of quiescent black holes are combined with our results for $\dot{M}$ and an estimate of the radiative efficiency of the accretion flow to conclude that quiescent black holes in the local universe are consistent with accreting in a locally thermally-stable state.  However, some fraction of galaxies may be recovering from a previous state thermal instability, either locally near the sphere of influence or as a part of much longer-term accretion events on galactic scales.  The maximum accretion rate of a thermally-stable flow is, however, too low to explain the accretion rates and short growth time of low mass black holes in the local universe.  The latter must instead result from gas being fed in from large radii, either due to galaxy mergers associated with hierarchical growth or thermal instabilities on larger (galactic) scales.  Our results have implications for topics ranging from the physical motivation for extrapolating the measured $L_{\rm X}-M_{\bullet}$ correlation to constrain the occupation fraction of SMBHs in low mass galaxies, to the diversity of circumnuclear media densities encountered by relativistic jets from tidal disruption events.
\end{abstract}
